---
<<<<<<< HEAD
title: "PA1_template.Rmd"
output: html_document
---
This is for Reproducible Research Course Project1

##1. Loading and preprocessing the data

```{r CACHE = TRUE}
setwd("/Users/wanhaochi/Desktop/Coursera/reproducible research/courseproject1")
data <- read.csv("activity.csv")
## take a first look at the data
summary(data)
str(data)

#pre-processing the data
data$seq <- rep(c(1:61),each = 288)
data$time <- rep(c(1:288),61)
data1 <- na.omit(data)
#change the format 
data1$interval0 <- formatC(data1$interval,width=4,format="d",flag="0")
data1$interval1 <-paste(data1$date,data$interval0)
data1$interval1 <- strptime(data1$interval1,"%Y-%m-%d %H%M")
data1$interval2 <- strptime(data1$interval0, "%H%M")

```

##2. What is mean total number of steps taken per day?

```{r CACHE = TRUE}
# Calculate the total number of steps taken each day
sum <- with(data1,
            tapply(X=steps,INDEX=seq,FUN=sum,simpilfy = T))

# Make a histogram of the total number of steps taken each day
hist(sum,xlab="Total number of steps",main = "Histogram of the total number of steps taken each day")
# Calculate and report the mean and median total number of steps taken each day
mean <- mean(sum)
median <- median(sum)
```

The mean total number of steps taken per day is `r mean`.
The median total number of steps taken per day is `r median`.

##3. What is the average daily activity pattern?

``` {r CACHE = TRUE}
mean_interval <- with(data1,
                      tapply(X=steps,INDEX=time,FUN=mean,simpilfy = T))

plot(unique(data1$interval2),mean_interval,type="l",xlab="Time",ylab="Steps",main="Averaged daily activity")

max <- which.max(mean_interval)
spot <- unique(data1$interval2)
x <- substr(spot[max],start =12, stop =19)

```

`r x` 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps.

##4.Imputting missing values

```{r CACHE = TRUE}
#Calculate and report the total number of missing values in the dataset
table(complete.cases(data))
```

The total number of missing values in the datasheet is 2304.

```{r CACHE = TRUE}
#Fill in all of the missing values with mean value of that 5-min interval in the datasheet.
data2 <- data.frame()
for (i in 1:288) {
    timesub <- subset(data,data$time == i)
    timesub$steps[is.na(timesub$steps)] <- mean_interval[i]
    data2 <- rbind(data2,timesub)
}

#double check the data after imputing the missing data
table(complete.cases(data2))

str(data2)

# Calculate the total number of steps taken each day in the new dataset
sum2 <- with(data2,
            tapply(X=steps,INDEX=seq,FUN=sum,simpilfy = T))

# Make a histogram of the total number of steps taken each day
hist(sum2,xlab="Total number of steps",main = "Histogram of the total number of steps taken each day")

# Calculate and report the mean and median total number of steps taken each day
mean2 <- mean(sum2)
median2 <- median(sum2)
```

The mean total number of steps taken per day after imputing the missing data is `r mean2` and before imputing the missing data is `r mean`.
The median total number of steps taken per day after imputing the missing data is `r median2` and before imputing the missing data is `r median`.

##5.Are there differences in activity patterns between weekdays and weekends?

```{r}

#change the format
data2$interval0 <- formatC(data2$interval,width=4,format="d",flag="0")
data2$interval <-paste(data2$date,data2$interval0)
data2$interval <- strptime(data2$interval,"%Y-%m-%d %H%M")
data2$interval1 <- strptime(data2$interval0, "%H%M")
data2$date <- strptime(data2$date, "%Y-%m-%d")
data2$day <- weekdays(data2$date)
weekday <- rep("weekday",5)
weekend <- rep("weekend",2)

#Create a new factor variable in the dataset with two levels
data2$weekday <- as.factor(c(rep(c(weekday,weekend),8),weekday))

# Calculate mean for weekday and weekend

data2_weekday <- subset(data2,data2$weekday == "weekday")
data2_weekend <- subset(data2,data2$weekday == "weekend")

mean_interval_weekday <- with(data2_weekday,
tapply(X=steps,INDEX=time,FUN=mean,simpilfy = T))

mean_interval_weekend <- with(data2_weekend,
tapply(X=steps,INDEX=time,FUN=mean,simpilfy = T))
#prepare data for plotting
mean_new <- data.frame(mean_interval_weekday,mean_interval_weekend,time = unique(data2$interval1))
mean_long <- with(mean_new,
                  rbind(data.frame(Day = "weekday",Time = time, Steps = mean_interval_weekday),
                        data.frame(Day = "weekend",Time=time,Steps = mean_interval_weekend))
)

#Plotting
library(ggplot2)
library(scales)
f1 = ggplot(data = mean_long, aes(x = Time, y = Steps,colour = Day) ) +
            geom_line() +
            facet_grid(Day ~ .)+
            ylab("Number of Steps")+
            scale_x_datetime(breaks=date_breaks("3 hour"),labels = date_format("%H:%M"))

print(f1)


=======
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
>>>>>>> 80edf39c3bb508fee88e3394542f967dd3fd3270
